---
title: "Hi"
output: html_notebook
---



```{r}

library(tidyverse)
library(arules)
library(dbplyr)
library(ggplot2)
library(backports)
library(reshape)

```

# Let's set the working directory to where our data is

```{r}

setwd("C:/Users/Decoded/Downloads/food-choices")


```

# Now let's read this in 

```{r}

master <- read.csv("food_coded.csv")
View(master)

```

# We need a method of creating a true false flag for the comfort reasons
# The grepl function is one we can use to convert string searches into YN flags

```{r}



?grepl()

```

# We can save the search for Boredom as a comfort rule reason

```{r}

master$boredom_check <- grepl('Boredom', master$comfort_food_reasons,  ignore.case = TRUE )

```


# we can repeat the same exercise for each of the other reasons

```{r}

master$boredom_check <- grepl('Boredom', master$comfort_food_reasons,  ignore.case = TRUE )

master$stress_check <- grepl('stress', master$comfort_food_reasons,  ignore.case = TRUE )

master$hunger_check <- grepl('hunger', master$comfort_food_reasons,  ignore.case = TRUE )

master$anger_check <- grepl('anger', master$comfort_food_reasons,  ignore.case = TRUE )


master$laziness_check <- grepl('laziness', master$comfort_food_reasons,  ignore.case = TRUE )

master$weather_check <- grepl('weather', master$comfort_food_reasons,  ignore.case = TRUE )

master$happiness_check <- grepl('happiness', master$comfort_food_reasons,  ignore.case = TRUE )

master$tv_check <- grepl('tv', master$comfort_food_reasons,  ignore.case = TRUE )

master$none_check <- grepl('none', master$comfort_food_reasons,  ignore.case = TRUE )


master$sad1 <- grepl('sad', master$comfort_food_reasons,  ignore.case = TRUE )
  
master$sad2 <-  grepl('depress', master$comfort_food_reasons,  ignore.case = TRUE )
master$sad_check <- FALSE
master$sad_check[master$sad1 == TRUE | master$sad2 == TRUE] <- TRUE
  
master$sad1 <- NULL

master$sad2 <- NULL

```



```{r}






```






# Now convert the comfort food items to lower case so that these are all lower case

# Then create a field in the transactions type so that we can perform some apriori association analysis on them

```{r}

master$comfort_food_list <- tolower(master$comfort_food)
master$reasons_list <-tolower(master$comfort_food_reasons)


new_ting <- paste(master$reasons_list, ", "  ,master$comfort_food_list)

?gsub

View(new_ting)
```

# apply to data frame not strings or vector of strings

# general function

```{r}

#master['comfort_food_list2'] <- lapply(master['comfort_food_list'],strsplit,",") #strsplit(master['comfort_food_list'], ",")

newtinga <- gsub(" / ", ", ", new_ting)

newtingb <- gsub("stres", "stress", newtinga)

newtingc <- gsub("lazyniss", "laziness", newtingb)

newtingd <- gsub("depression", "sadness", newtingc)

new_ting2 <- gsub("  ", " ", newtingd)

new_ting3 <- gsub("mac and cheese", "mac n cheese", new_ting2)

new_ting4 <- gsub(" and ", ", ", new_ting3)

new_ting5 <- gsub("\\.", "", new_ting4)

new_ting6 <- gsub(" ,", ",",  new_ting5)

new_ting7 <- gsub(",,", ",",  new_ting6)

new_ting8 <- gsub(" ,", ",",  new_ting7)

new_ting9 <- gsub("they taste better than other food they are a pickme up they are easy to make", "taste", new_ting8)

new_ting10 <- gsub("grandma homemade chocolate cake anything homemade", "cake", new_ting9)

new_ting11 <-   gsub("i don/'t eat comfort food i just eat when i'm hungry", "hunger", new_ting10)



second_thing <- strsplit(new_ting11 , ", ")

 third_thing <- sapply(second_thing, trimws )

baskets <- as(third_thing, "transactions")

inspect(baskets[1:10])
```


```{r}

rules = apriori(data=baskets,parameter=list(support=0.001, confidence=0.3, maxlen=10))

```

```{r}

inspect(rules[1:30])


```




```{r}

stress_rules = sort(
  subset(rules, subset=lhs %in% 'stress'),
  by = 'lift',decreasing = T
)

inspect(stress_rules[1:20])

```

```{r}
stress_rules2 = DATAFRAME(stress_rules)

stress_rules2[stress_rules2['LHS'] == "{stress}", ]
```


```{r}

sadness_rules = sort(
  subset(rules, subset=lhs %in% 'sadness'),
  by = 'lift',decreasing = T
)
sadness_rules2 = DATAFRAME(sadness_rules)
#sadder_rules = subset(sad_rules, subset=!(rhs %in% 'excited')


sadness_rules2[sadness_rules2['LHS'] == "{sadness}", ]
```


#IF condition
# dataset$thing[dataset$marital_status == 2] <- 15
